
```{r include=FALSE}
# Set working directory. Change as needed
setwd('C:/Users/a6p/Desktop/Uni 2014/E1161 - Collaborative Research/Final Project/')
```

# Methodology

We obtain the patent data on the city level from the OECD database. First we install the R [@CiteR] package [@RSDMX] necessary for obtaining the OECD dataset.

```{r echo=FALSE}
# Install packages to pull OECD data

# install.packages("devtools")
require("devtools")
library("rsdmx")
```

In the next step we pull the OECD data from the online database using [@RSDMX]. As we are interested in patent data, we work with data indicating the PCT patent applications per 10,000 inhabitants. As control variables we also use GDP per capita and green space available per inhabitant in square meters per inhabitant.
, y

```{r include=FALSE}
# Pull OECD city patent data
dataURL <- "http://stats.oecd.org/restsdmx/sdmx.ashx/GetData/CITIES/DEU+DE001+DE002+DE003+DE007+DE035+DE507+ESP+ES001+ES002+FRA+FR001+ITA+IT001+IT002+IT009+SWE+SE001+SE002+SE003+USA+US014+US048+US084+US146+US234.PCT_INTENSITY?startTime=2000&endTime=2012"

sdmx <- readSDMX(dataURL)
city_pat <- as.data.frame(sdmx)

# Clean table by deleting and renaming rows
city_pat$VAR <- NULL # Delete columns
city_pat$attrs.df <- NULL

colnames(city_pat) <- c("METRO_ID" , "Year" , "Patent_Intensity") # Rename columns

# Pull OECD GDP per capita data

dataURL2 <- "http://stats.oecd.org/restsdmx/sdmx.ashx/GetData/CITIES/DEU+DE001+DE002+DE003+DE007+DE035+DE507+ESP+ES001+ES002+FRA+FR001+ITA+IT001+IT002+IT009+SWE+SE001+SE002+SE003+USA+US014+US048+US084+US146+US234.GDP_PC?startTime=2000&endTime=2012"

sdmx <- readSDMX(dataURL2)
city_gdp <- as.data.frame(sdmx)

# Clean table by deleting and renaming rows
city_gdp$VAR <- NULL # Delete columns
city_gdp$attrs.df <- NULL

colnames(city_gdp) <- c("METRO_ID" , "Year" , "GDP") # Rename columns

# Pull OECD environmental data

dataURL3 <- "http://stats.oecd.org/restsdmx/sdmx.ashx/GetData/CITIES/DEU+DE001+DE002+DE003+DE007+DE035+DE507+ESP+ES001+ES002+FRA+FR001+ITA+IT001+IT002+IT009+SWE+SE001+SE002+SE003+USA+US014+US048+US084+US146+US234.GREEN_AREA_PC?startTime=2000&endTime=2008"

sdmx <- readSDMX(dataURL3)
city_env <- as.data.frame(sdmx)

# Clean table by deleting and renaming rows
city_env$VAR <- NULL # Delete columns
city_env$attrs.df <- NULL

colnames(city_env) <- c("METRO_ID" , "Year" , "GreenArea") # Rename columns

```

After cleaning the data frames and sorting the data, we merge the three data sets into one using the following code: 


```{r echo=TRUE}
# Merge all 3 OECD citydata sets

oecd_pat_gdp <- merge(city_pat , city_gdp , by=c("METRO_ID" , "Year"))
oecd <- merge(oecd_pat_gdp , city_env , by=c("METRO_ID" , "Year"))

```

To further clean up the table we rename the city and country IDs.

```{r echo=FALSE}

# Clean city table, insert city and country names
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DEU" , "Germany")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE001" , "Berlin")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE002" , "Hamburg")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE003" , "Munich")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE007" , "Stuttgart")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE035" , "Karlsruhe")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "DE507" , "Aachen")

oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "ESP" , "Spain")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "ES001" , "Madrid")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "ES002" , "Barcelona")

oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "FRA" , "France")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "FR001" , "Paris")

oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "ITA" , "Italy")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "IT001" , "Rome")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "IT002" , "Milan")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "IT009" , "Bologna")

oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "SWE" , "Sweden")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "SE001" , "Stockholm")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "SE002" , "Gothenburg")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "SE003" , "Malmö")

oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "USA" , "United States")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "US014" , "Minneapolis")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "US048" , "Boston")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "US084" , "New York")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "US146" , "San Francisco")
oecd$METRO_ID <- replace(oecd$METRO_ID , oecd$METRO_ID == "US234" , "Austin")

```

# Descriptive statistics

The characteristics of the city data obtained through the OECD database are shown in this section.

In order to find the mean of the patent data, we create a function ($\bar{x} = \frac{\sum x}{n}$) of the vector.

```{r}
fun_mean <- function(x){
    sum(x) / length(x)
}
mean(oecd$Patent_Intensity)
```

In the data set, the mean for PCT patent application per 10,000 inhabitants is 3.44. The histogram indicates that the mean is not very meaningful. There are many cities with patent applications much lower and some cities with around 5.5 patent applications. This finding supports our claim that on the regional level there are great variations in patent activity, which undermines national level analyses.


```{r, include=TRUE}
hist(oecd$Patent_Intensity,
      main = 'PCT applications',
      xlab = 'PCT patent applications per 10,000 inhabitants')
```

The histogram indicates the relationship between GDP and patent intensity. 

```{r, include=TRUE}
plot(oecd$Patent_Intensity , oecd$GDP, 
      main = 'Patent intensity and GDP per capita',
      xlab = 'PCT patent applications per 10,000 inhabitants',
      ylab = 'GDP per capita')
```

A second histogram shows the relationship between green space available and patent intensity. The relationship seems stronger than with GDP.


```{r, include=TRUE}
plot(oecd$Patent_Intensity , oecd$Grenarea, 
      main = 'Patent intensity and GDP per capita',
      xlab = 'PCT patent applications per 10,000 inhabitants',
      ylab = 'Green space in square meters per inhabitant')
```

